quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-05-09 paper-quollr.Rmd
Added hexagonal grid with the notifications

>>>>

```{r}
#| label: code-illustration
# Code to draw illustration for notation
## hexagon binning to have regular hexagons
hb_obj_notation <- hex_binning(
  nldr_obj = scurve_umap_obj, 
  bin1 = 9, 
  q = 0.1)

a1_temp <- hb_obj_notation$a1
a2_temp <- hb_obj_notation$a2
l_temp <- quad(a=3, b = 2 * a2_temp, c = -(a2_temp^2 + a1_temp^2))

## Data set with all centroids
all_centroids_df_temp <- hb_obj_notation$centroids
hex_grid_temp <- hb_obj_notation$hex_poly

hex_grid_temp45 <- hex_grid_temp |> 
  filter(hex_poly_id == 45)

start_pt <- all_centroids_df_temp |> 
  filter(hexID == 1)
d_rect <- tibble(x1min = 0, 
                 x1max = 1,
                 x2min = 0,
                 x2max = diff(scurve_umap_obj$lim2)/diff(scurve_umap_obj$lim1)) # x2max = r2

# To move the rectangle to ignore the overlap with the centroids
# rect_adj <- tibble(x1 = 0.03, x2 = 0.03)
rect_adj <- tibble(x1 = -0.03, x2 = 0.03)

a1 <- tibble(x = all_centroids_df_temp$c_x[4],
             xend = all_centroids_df_temp$c_x[5],
             y = all_centroids_df_temp$c_y[21],
             yend = all_centroids_df_temp$c_y[21],
             label = expression(a[1]))
a2 <- tibble(x = all_centroids_df_temp$c_x[25],
             xend = all_centroids_df_temp$c_x[25],
             y = all_centroids_df_temp$c_y[25],
             yend = all_centroids_df_temp$c_y[33],
             label = expression(a[2]))
l <- tibble(x = hex_grid_temp45$x[2],
            xend = hex_grid_temp45$x[3],
            y = hex_grid_temp45$y[2],
            yend = hex_grid_temp45$y[3],
            label = expression(l))

hex_param_vis <- ggplot() + 
    geom_polygon(data = hex_grid_temp, 
                        aes(x = x, 
                            y = y, 
                            group = hex_poly_id),
                 fill = "white", 
                 color = "#bdbdbd") +
    geom_point(data = all_centroids_df_temp, aes(
      x = c_x, 
      y = c_y), 
      color = "#31a354", size = 0.9) +
    geom_point(data = start_pt, aes(x = c_x, 
                                    y = c_y), 
               color = "black") + 
    geom_rect(data=d_rect, 
              aes(xmin = x1min - rect_adj$x1,# - rect_adj$s1, 
                  xmax = x1max - rect_adj$x1,# - rect_adj$s1, 
                  ymin = x2min - rect_adj$x2,# - rect_adj$s2, 
                  ymax = x2max - rect_adj$x2),# - rect_adj$s2), 
              fill = "white", 
              color = "black", 
              alpha = 0, 
              linewidth = 0.7) +
    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
                                y=x2min - rect_adj$x2)) + 
    geom_point(data=d_rect, aes(x=x1max - rect_adj$x1, 
                                y=x2min - rect_adj$x2)) + 
    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
                                y=x2max - rect_adj$x2)) + 
    annotate("text", x=d_rect$x1min - rect_adj$x1, 
                     y=d_rect$x2min - rect_adj$x2,
                     label = "(0,0)", 
             hjust=-0.1, vjust=-0.3) + 
    annotate("text", x=d_rect$x1max - rect_adj$x1, 
                     y=d_rect$x2min - rect_adj$x2,
                     label = "(0,1)", 
             hjust=1.1, vjust=-0.3) + 
    annotate("text", x=d_rect$x1min - rect_adj$x1, 
                     y=d_rect$x2max - rect_adj$x2,
                     label = expression(group("(", 
                        list(0, y[2][max]),")")), 
            hjust=-0.1, vjust=1.2) + 
    geom_segment(data=d_rect, aes(
      x = x1min  - rect_adj$x1, # 0 - 0.03, 
      y = -0.31, 
      xend = x1max - rect_adj$x1, #1 - 0.03, 
      yend = -0.31), #-0.35),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
                 color = "black")+
    annotate("text", x=0.5, y=-0.36, 
             label = expression(r[1]), color = "black") +
    geom_segment(data=d_rect, aes(
      x = -0.25, 
      y = x2min - rect_adj$x2, #0 - 0.05, 
      xend = -0.25, 
      yend = x2max - rect_adj$x2), #r2 - 0.05),
      arrow = arrow(length = unit(0.03, "npc"),
                       ends = "both"), 
                 color = "black")+ 
    annotate("text", x=-0.3, y=0.4, 
             label = expression(r[2]), color = "black") +
    geom_segment(data = a1, aes(
      x = x, #-0.1 + 0.2087578, 
      y = y, #-0.15, 
      xend = xend, #-0.1 + 0.2087578*2, 
      yend = yend), #-0.15),
      arrow = arrow(length = unit(0.03, "npc"),
        ends = "both"), 
        color = "black")+ # a1 = 0.2087578
    annotate("text", 
             x=(a1$x+a1$xend)/2, 
             y=a1$y, 
             label = expression(a[1]), 
             color = "black",
             vjust = 1.2) +
    geom_segment(data = a2, aes(
      x = x, #-0.15, 
      y = y, #-0.1*r2 + 0.1807896*2, 
      xend = xend, #-0.15, 
      yend = yend), #-0.1*r2 + 0.1807896*3),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
      color = "black") + # a2 = 0.1807896
    annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
             label = expression(a[2]), 
             color = "black", hjust=-0.2) +
    annotate("text", x=-0.18, y=-0.24, 
      label = expression(group("(", list(s[1], s[2]), ")")),
      color = "black") +
  geom_segment(data = l, aes(
      x = x, #-0.15, 
      y = y, #-0.1*r2 + 0.1807896*2, 
      xend = xend, #-0.15, 
      yend = yend), #-0.1*r2 + 0.1807896*3),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
      color = "black") + 
    annotate("text", x=l$x + 0.03, y=(l$y+l$yend)/2, 
             label = expression(l), 
             color = "black", hjust=-0.2) +
  coord_equal()

```{r}
#| label: fig-hex-param
#| fig-cap: "The components of the hexagon grid illustrating notation."

hex_param_vis
```
